---
title: "Project : Propensity Score on Yahoo Personals Data 2004-2005, M. Sherrod"
output: pdf_document
---


## R Packages
You will need to use an R package that you may not already have installed, arm. 


```{r setup, include=FALSE, echo=F}
setwd("~/R")
source("OG_YH_cleaningv_NAImputev2.r", local = knitr::knit_global())
#Marlee Sherrod

library(arm)
library(tidyverse)
library(cobalt)
library(MatchIt)
library(sandwich)
library(lmtest)
library(Hmisc)
library(MASS)


```



#### Load the data and choose confounders (5 points)

```{r}
# load data
YHdatNA <- OG_YH_test
colnames(YHdatNA)
summary(YHdatNA$CGRAD)

sum(YHdatNA$CGRAD == 1) #3462
sum(YHdatNA$CGRAD == 0) #2408
```


```{r}
# print out the names of all your confounders
YHdatNA_NAMES <- YHdatNA %>% dplyr::select(-c(1, 21:22, 44:209)) #removing ID column, EDUCATION AND CGRAD COLUMNS, and D_Columns

confounder_names <- colnames(YHdatNA_NAMES)
confounder_names
```

#Plotting Data
```{r}
#plotting counts of cGrad by region
pcgradreg <- ggplot(data = YHdatNA, aes(x = as.factor(CGRAD), fill = REGION)) + 
    geom_bar() +
    labs(title = "No. College Grads and Non-Grads by Region",  # Adding title
       x = "College Graduate",  # X-axis label
       y = "Count") +
  geom_text(stat = 'count', aes(label = ..count..), position = position_stack(vjust = 0.5)) 
pcgradreg

```


```{r}
pcgradwant_kids <- ggplot(data = YHdatNA, aes(x = as.factor(CGRAD), fill = D_WANTKIDS_YES)) +
    geom_bar() +
      labs(title = "Non-C. Grads and C .Grads w/ 'Does Not Want Kids' Preference in Partner",  # Adding title
       x = "College Graduate",  # X-axis label
       y = "Count") +
  geom_text(stat = 'count', aes(label = ..count..), position = position_stack(vjust = 0.5)) 
pcgradwant_kids

#ggsave("pcgradwant_kids.png", pcgradwant_kids, width = 6, height = 4, dpi = 300)
#dev.off()
```

```{r}
pcgradsex <- ggplot(data = YHdatNA, aes(x = as.factor(CGRAD), fill = SEX)) +
    geom_bar() +
      labs(title = "No. College Grads and Non-Grads by Sex",  # Adding title
       x = "College Graduate",  # X-axis label
       y = "Count") +
  geom_text(stat = 'count', aes(label = ..count..), position = position_stack(vjust = 0.5)) 

pcgradsex
```
 

```{r}
pcgradwant_kids_region <- ggplot(data = YHdatNA, aes(x = as.factor(CGRAD), fill = D_WANTKIDS_YES)) +
    geom_bar() +
  facet_wrap(~ REGION) +
      labs(title = "Non-C.Grads/C.Grads w. 'Does Not Want Kids' Preference in Partner by Region",  # Adding title
       x = "College Graduate",  # X-axis label
       y = "Count") +
  geom_text(stat = 'count', aes(label = ..count..), position = position_stack(vjust = 0.5)) 
pcgradwant_kids_region


```




#### Step 2: Look at original balance and overlap


#PROBIT MODELS

```{r}
bal_ov_B4_prob_ATC <- matchit(CGRAD ~ SEX + REGION + AGE + ETHNICITY + MARSTAT + BODY + HEIGHT_FT + HEIGHT_IN + SMOKING + DRINKING + LIVE_ALONE + LIVE_KIDS + LIVE_PAR + LIVE_PETS + LIVE_ROOM + LIVE_FAM + LIVE_PARTY + HAVEKIDS + WANTKIDS + EMPSTAT + OCCUP + INCOME + RELIGION + SERVICES + POLITICS + LANG_ENGLISH + LANG_FRENCH + LANG_GERMAN + LANG_ITALIAN + LANG_SPANISH + LANG_PORTUGUESE + LANG_DUTCH + LANG_CHINESE + LANG_JAPANESE + LANG_ARABIC + LANG_RUSSIAN + LANG_HEBREW + LANG_HINDI + LANG_TAGALOG + LANG_URDU, data = YHdatNA,
                 method = NULL, distance = "glm", link = "probit", estimand = "ATC")

bal_ov_B4_prob_ATC
sum_bal4_ATC <- summary(bal_ov_B4_prob_ATC)

summarydf_sum_bal4_ATC <- cbind(
  "Means Treated B4" = sum_bal4_ATC$sum.all[,"Means Treated"],
  "Means Treated M" = sum_bal4_ATC$sum.matched[,"Means Treated"],
  "Means Control B4" = sum_bal4_ATC$sum.all[,"Means Control"],
  "Means Control M" = sum_bal4_ATC$sum.matched[,"Means Control"],
  "Std. Mean Diff. B4" = sum_bal4_ATC$sum.all[,"Std. Mean Diff."],
  "Std. Mean Diff. M" = sum_bal4_ATC$sum.matched[,"Std. Mean Diff."],
  "Var. Ratio B4" = sum_bal4_ATC$sum.all[,"Var. Ratio"],
  "Var. Ratio M" = sum_bal4_ATC$sum.matched[,"Var. Ratio"],
  "eCDF Mean B4" = sum_bal4_ATC$sum.all[,"eCDF Mean"],
  "eCDF Mean M" = sum_bal4_ATC$sum.matched[,"eCDF Mean"],
  "eCDF Max B4" = sum_bal4_ATC$sum.all[,"eCDF Max"],
  "eCDF Max M" = sum_bal4_ATC$sum.matched[,"eCDF Max"],
  "Std. Pair Dist. B4" = sum_bal4_ATC$sum.all[,"Std. Pair Dist."],
  "Std. Pair Dist. M" = sum_bal4_ATC$sum.matched[,"Std. Pair Dist."]
) 

summarydf_sum_bal4_ATC <- as.data.frame(summarydf_sum_bal4_ATC)

view(summarydf_sum_bal4_ATC)

summarynn_sum_bal4_ATC <- sum_bal4_ATC$nn
summarynn_sum_bal4_ATC

bal.plot(bal_ov_B4_prob_ATC, type = "histogram") #plot looks pretty good


stdm_summarydf_sum_bal4_ATC <- summarydf_sum_bal4_ATC[order(-abs(summarydf_sum_bal4_ATC$"Std. Mean Diff. B4")), ] #sorting it so its in descending order (largest std to smallest)
table1 <- round(head(stdm_summarydf_sum_bal4_ATC, n= 7), digits = 4)
write.csv(table1, "table1.csv", row.names = T)

kableExtra::kable(stdm_summarydf_sum_bal4_ATC)
```

#### Step 3: Create a propensity model and assess balance and overlap - NEW (to get absolute mean difference. We eant to use eCDF Mean
 
```{r}
mat_nearest_ATC_wr <- matchit(CGRAD ~ SEX + REGION + AGE + ETHNICITY + MARSTAT + BODY + HEIGHT_FT + HEIGHT_IN + SMOKING + DRINKING + LIVE_ALONE + LIVE_KIDS + LIVE_PAR + LIVE_PETS + LIVE_ROOM + LIVE_FAM + LIVE_PARTY + HAVEKIDS + WANTKIDS + EMPSTAT + OCCUP + INCOME + RELIGION + SERVICES + POLITICS + LANG_ENGLISH + LANG_FRENCH + LANG_GERMAN + LANG_ITALIAN + LANG_SPANISH + LANG_PORTUGUESE + LANG_DUTCH + LANG_CHINESE + LANG_JAPANESE + LANG_ARABIC + LANG_RUSSIAN + LANG_HEBREW + LANG_HINDI + LANG_TAGALOG + LANG_URDU, data = YHdatNA, method = "nearest", estimand = "ATC", distance = "glm", link = "probit", replace = T) 
                        

mat_nearest_ATC_wr
options(max.print=3000)
sum_MatwrATC <- summary(mat_nearest_ATC_wr)
#sum_fullATE

summarydf_sum_MatwrATC <- cbind(
  "Means Treated B4" = sum_MatwrATC$sum.all[,"Means Treated"],
  "Means Treated M" = sum_MatwrATC$sum.matched[,"Means Treated"],
  "Means Control B4" = sum_MatwrATC$sum.all[,"Means Control"],
  "Means Control M" = sum_MatwrATC$sum.matched[,"Means Control"],
  "Std. Mean Diff. B4" = sum_MatwrATC$sum.all[,"Std. Mean Diff."],
  "Std. Mean Diff. M" = sum_MatwrATC$sum.matched[,"Std. Mean Diff."],
  "Var. Ratio B4" = sum_MatwrATC$sum.all[,"Var. Ratio"],
  "Var. Ratio M" = sum_MatwrATC$sum.matched[,"Var. Ratio"],
  "eCDF Mean B4" = sum_MatwrATC$sum.all[,"eCDF Mean"],
  "eCDF Mean M" = sum_MatwrATC$sum.matched[,"eCDF Mean"],
  "eCDF Max B4" = sum_MatwrATC$sum.all[,"eCDF Max"],
  "eCDF Max M" = sum_MatwrATC$sum.matched[,"eCDF Max"],
  "Std. Pair Dist. B4" = sum_MatwrATC$sum.all[,"Std. Pair Dist."],
  "Std. Pair Dist. M" = sum_MatwrATC$sum.matched[,"Std. Pair Dist."]
)

summarydf_sum_MatwrATC <- as.data.frame(summarydf_sum_MatwrATC)

view(summarydf_sum_MatwrATC) 

summarynn_sum_MatwrATC <- sum_MatwrATC$nn
summarynn_sum_MatwrATC


bal.plot(mat_nearest_ATC_wr, type = "histogram") 

stdm_sum_MatwrATC <- summarydf_sum_MatwrATC[order(-abs(summarydf_sum_MatwrATC$"eCDF Mean M")), ] #sorting it so its in descending order (largest absolute mean difference to smallest)

stdm_sum_MatwrATC$actualdiff <- stdm_sum_MatwrATC$`Means Control M` - stdm_sum_MatwrATC$`Means Treated M`

#check that the actual difference is equal to eCDF Mean M
stdm_sum_MatwrATC$actualdiff <- abs(stdm_sum_MatwrATC$`Means Control M` - stdm_sum_MatwrATC$`Means Treated M`)

#check and see which variables where they don't match
view(stdm_sum_MatwrATC %>% filter(`eCDF Mean M` != actualdiff))
#they are continuous variables and their standard differences are under 0.10


 #sorting it so its in descending order (largest absval to smallest)
table2 <- round(head(stdm_sum_MatwrATC, n= 7), digits = 4)
setwd("~/R")
write.csv(table2, "table2.csv", row.names = T)


#sort covariates by absolute value of mean difference
#stdm_sum_MatwrATC <- stdm_sum_MatwrATC %>% relocate("eCDF Mean M", .before = "Means Treated B4") #moving std means column closer

kableExtra::kable(stdm_sum_MatwrATC)
```

#### Step 3: Create a propensity model and assess balance and overlap

```{r}
# 
# 
# mat_nearest_ATC_wr <- matchit(CGRAD ~ SEX + REGION + AGE + ETHNICITY + MARSTAT + BODY + HEIGHT_FT + HEIGHT_IN + SMOKING + DRINKING + LIVE_ALONE + LIVE_KIDS + LIVE_PAR + LIVE_PETS + LIVE_ROOM + LIVE_FAM + LIVE_PARTY + HAVEKIDS + WANTKIDS + EMPSTAT + OCCUP + INCOME + RELIGION + SERVICES + POLITICS + LANG_ENGLISH + LANG_FRENCH + LANG_GERMAN + LANG_ITALIAN + LANG_SPANISH + LANG_PORTUGUESE + LANG_DUTCH + LANG_CHINESE + LANG_JAPANESE + LANG_ARABIC + LANG_RUSSIAN + LANG_HEBREW + LANG_HINDI + LANG_TAGALOG + LANG_URDU, data = YHdatNA, method = "nearest", estimand = "ATC", distance = "glm", link = "probit", replace = T) 
#                         
# 
# mat_nearest_ATC_wr
# options(max.print=3000)
# sum_MatwrATC <- summary(mat_nearest_ATC_wr)
# #sum_fullATE
# 
# summarydf_sum_MatwrATC <- cbind(
#   "Means Treated B4" = sum_MatwrATC$sum.all[,"Means Treated"],
#   "Means Treated M" = sum_MatwrATC$sum.matched[,"Means Treated"],
#   "Means Control B4" = sum_MatwrATC$sum.all[,"Means Control"],
#   "Means Control M" = sum_MatwrATC$sum.matched[,"Means Control"],
#   "Std. Mean Diff. B4" = sum_MatwrATC$sum.all[,"Std. Mean Diff."],
#   "Std. Mean Diff. M" = sum_MatwrATC$sum.matched[,"Std. Mean Diff."],
#   "Var. Ratio B4" = sum_MatwrATC$sum.all[,"Var. Ratio"],
#   "Var. Ratio M" = sum_MatwrATC$sum.matched[,"Var. Ratio"],
#   "eCDF Mean B4" = sum_MatwrATC$sum.all[,"eCDF Mean"],
#   "eCDF Mean M" = sum_MatwrATC$sum.matched[,"eCDF Mean"],
#   "eCDF Max B4" = sum_MatwrATC$sum.all[,"eCDF Max"],
#   "eCDF Max M" = sum_MatwrATC$sum.matched[,"eCDF Max"],
#   "Std. Pair Dist. B4" = sum_MatwrATC$sum.all[,"Std. Pair Dist."],
#   "Std. Pair Dist. M" = sum_MatwrATC$sum.matched[,"Std. Pair Dist."]
# )
# 
# summarydf_sum_MatwrATC <- as.data.frame(summarydf_sum_MatwrATC)
# 
# view(summarydf_sum_MatwrATC) 
# 
# summarynn_sum_MatwrATC <- sum_MatwrATC$nn
# summarynn_sum_MatwrATC
# 
# 
# bal.plot(mat_nearest_ATC_wr, type = "histogram") 
# 
# stdm_sum_MatwrATC <- summarydf_sum_MatwrATC[order(-abs(summarydf_sum_MatwrATC$"Std. Mean Diff. M")), ] #sorting it so its in descending order (largest std to smallest)
# 
#  #sorting it so its in descending order (largest std to smallest)
# table2 <- round(head(stdm_sum_MatwrATC, n= 7), digits = 4)
# setwd("~/R")
# write.csv(table2, "table2.csv", row.names = T)
# 
# 
# #sort covariates by absolute value of standarzied mean difference
# stdm_sum_MatwrATC <- stdm_sum_MatwrATC %>% relocate("Std. Mean Diff. M", .before = "Means Treated B4") #moving std means column closer
# 
# stdm_sum_MatwrATC

```

#### Step 4: Edit Propensity Score model to create better balance

```{r}
#Final Model
#just interaction between Havekids and wantkids

mat_nearest_ATC_wr4 <- matchit(CGRAD ~ SEX + REGION + AGE + ETHNICITY + MARSTAT + BODY + HEIGHT_FT + HEIGHT_IN + SMOKING + DRINKING + LIVE_ALONE + LIVE_KIDS + LIVE_PAR + LIVE_PETS + LIVE_ROOM + LIVE_FAM + LIVE_PARTY + HAVEKIDS*WANTKIDS + EMPSTAT + OCCUP + INCOME + RELIGION + SERVICES + POLITICS + LANG_ENGLISH + LANG_FRENCH + LANG_GERMAN + LANG_ITALIAN + LANG_SPANISH + LANG_PORTUGUESE + LANG_DUTCH + LANG_CHINESE + LANG_JAPANESE + LANG_ARABIC + LANG_RUSSIAN + LANG_HEBREW + LANG_HINDI + LANG_TAGALOG + LANG_URDU, data = YHdatNA,
                 method = "nearest", estimand = "ATC", distance = "glm", link = "probit", replace = T) 

mat_nearest_ATC_wr4
options(max.print=3000)
sum_MatwrATC4 <- summary(mat_nearest_ATC_wr4)
#sum_MatwrATC4


summarydf_sum_MatwrATC4 <- cbind(
  "Means Treated B4" = sum_MatwrATC4$sum.all[,"Means Treated"],
  "Means Treated M" = sum_MatwrATC4$sum.matched[,"Means Treated"],
  "Means Control B4" = sum_MatwrATC4$sum.all[,"Means Control"],
  "Means Control M" = sum_MatwrATC4$sum.matched[,"Means Control"],
  "Std. Mean Diff. B4" = sum_MatwrATC4$sum.all[,"Std. Mean Diff."],
  "Std. Mean Diff. M" = sum_MatwrATC4$sum.matched[,"Std. Mean Diff."],
  "Var. Ratio B4" = sum_MatwrATC4$sum.all[,"Var. Ratio"],
  "Var. Ratio M" = sum_MatwrATC4$sum.matched[,"Var. Ratio"],
  "eCDF Mean B4" = sum_MatwrATC4$sum.all[,"eCDF Mean"],
  "eCDF Mean M" = sum_MatwrATC4$sum.matched[,"eCDF Mean"],
  "eCDF Max B4" = sum_MatwrATC4$sum.all[,"eCDF Max"],
  "eCDF Max M" = sum_MatwrATC4$sum.matched[,"eCDF Max"],
  "Std. Pair Dist. B4" = sum_MatwrATC4$sum.all[,"Std. Pair Dist."],
  "Std. Pair Dist. M" = sum_MatwrATC4$sum.matched[,"Std. Pair Dist."]
) 

summarydf_sum_MatwrATC4 <- as.data.frame(summarydf_sum_MatwrATC4)

view(summarydf_sum_MatwrATC4)

summarynn_sum_MatwrATC4 <- sum_MatwrATC4$nn
summarynn_sum_MatwrATC4

bal.plot(mat_nearest_ATC_wr4, type = "histogram") 


#sort covariates by absolute value mean difference

stdm_sum_MatwrATC4 <- summarydf_sum_MatwrATC4[order(-abs(summarydf_sum_MatwrATC4$"eCDF Mean M")), ] #sorting it so its in descending order (largest std to smallest)

# table3 <- round(head(stdm_sum_MatwrATC4, n= 7), digits = 4)
# setwd("~/R")
# write.csv(table3, "table3.csv", row.names = T)

stdm_sum_MatwrATC4 <- stdm_sum_MatwrATC4 %>% relocate("eCDF Mean M", .before = "Means Treated B4") #moving std means column closer
stdm_sum_MatwrATC4

```


#### Step 5: Estimate Treatment Effects

```{r}
#get the weights of all the observations
weights_ATC_wr4 <- as.vector(mat_nearest_ATC_wr4$weights)

YHdatNA$D_WANTKIDS_NO_v2 <- ifelse(YHdatNA$D_WANTKIDS_NO == "(1) yes", 1, 0) #yes, they want't to date someone who doesn't want kids


#Warning: non-integer #successes in a binomial glm!
Mat_Mod_ATC_wr4_log <- glm(D_WANTKIDS_NO_v2  ~ as.factor(CGRAD) + REGION + as.factor(SEX) + AGE +  ETHNICITY + MARSTAT + BODY + HEIGHT_FT + HEIGHT_IN + SMOKING + DRINKING + LIVE_ALONE + LIVE_KIDS + LIVE_PAR + LIVE_PETS + LIVE_ROOM + LIVE_FAM + LIVE_PARTY + HAVEKIDS + WANTKIDS + EMPSTAT + OCCUP + INCOME + RELIGION + SERVICES + POLITICS + LANG_ENGLISH + LANG_FRENCH + LANG_GERMAN + LANG_ITALIAN + LANG_SPANISH + LANG_PORTUGUESE + LANG_DUTCH + LANG_CHINESE + LANG_JAPANESE + LANG_ARABIC + LANG_RUSSIAN + LANG_HEBREW + LANG_HINDI + LANG_TAGALOG + LANG_URDU, data = YHdatNA, weights = weights_ATC_wr4, family = binomial("logit"))


options(max.print=50)
Mat_Mod_ATC_wr_summary <- summary(Mat_Mod_ATC_wr4_log)

sink("lm.txt")
print(Mat_Mod_ATC_wr_summary) 
sink()


point_estimate_matchATC_wr4_log <- (coeftest(Mat_Mod_ATC_wr4_log, vcov = sandwich)[2,1])
point_estimate_matchATC_wr4_log

probcoef <- point_estimate_matchATC_wr4_log/4
probcoef

point_SE_matchATC_wr4_log <- (coeftest(Mat_Mod_ATC_wr4_log, vcov = sandwich)[2,2])
point_SE_matchATC_wr4_log

```

```{r}
#linear regression for final compare

Mat_Mod_ATC_wr4_lm_vanilla <- lm(D_WANTKIDS_NO_v2 ~ as.factor(CGRAD) + REGION + as.factor(SEX) + AGE +  ETHNICITY + MARSTAT + BODY + HEIGHT_FT + HEIGHT_IN + SMOKING + DRINKING + LIVE_ALONE + LIVE_KIDS + LIVE_PAR + LIVE_PETS + LIVE_ROOM + LIVE_FAM + LIVE_PARTY + HAVEKIDS + WANTKIDS + EMPSTAT + OCCUP + INCOME + RELIGION + SERVICES + POLITICS + LANG_ENGLISH + LANG_FRENCH + LANG_GERMAN + LANG_ITALIAN + LANG_SPANISH + LANG_PORTUGUESE + LANG_DUTCH + LANG_CHINESE + LANG_JAPANESE + LANG_ARABIC + LANG_RUSSIAN + LANG_HEBREW + LANG_HINDI + LANG_TAGALOG + LANG_URDU, data = YHdatNA)

options(max.print=50)
Mat_Mod_ATC_wr4_lm_vanilla_summary <- summary(Mat_Mod_ATC_wr4_lm_vanilla)
Mat_Mod_ATC_wr4_lm_vanilla_summary

sink("lm_van.txt")
print(Mat_Mod_ATC_wr4_lm_vanilla_summary)
sink()

point_estimate_matchATC_wr4_lm_van<- (coeftest(Mat_Mod_ATC_wr4_lm_vanilla, vcov = sandwich)[2,1])
point_estimate_matchATC_wr4_lm_van

point_SE_matchATC_wr4_lm_van <- (coeftest(Mat_Mod_ATC_wr4_lm_vanilla, vcov = sandwich)[2,2])
point_SE_matchATC_wr4_lm_van



```

